---
title:    "Two-way Factorial ANOVA"
subtitle: "Interactions"
author:   "Dr. Sarah Schwartz"
date:     "USU: Educ/Psy 6600"
output: pdf_document
---

```{r oppts, include=FALSE}
# set global chunk options...  
#  this changes the defaults so you don't have to repeat yourself
knitr::opts_chunk$set(comment     = NA,
                      cache       = TRUE,
                      echo        = TRUE, 
                      warning     = FALSE, 
                      message     = FALSE,
                      error       = FALSE,
                      fig.align   = "center", # center all figures
                      fig.width   = 5,        # set default figure width to 5 inches
                      fig.height  = 3)        # set default figure height to 3 inches

```

# Preparation

## Packages

```{r}
library(tidyverse)
library(furniture)
library(afex)
library(emmeans)
```

\clearpage

## Functions

### Data Generation  

* Row Factor = `drug`, 2 levels
    1. `Antidepressant`
    2. `Placebo`


* Column Factor = `therapy`, 3 levels
    1. `Psychoanalysis`
    2. `Group Therapy`
    3. `Behavior Modification`


Function Options:
* `mu_j` = cell means vector, *default = 2*
* `n_j` = cell sub-sample size vector, *default = 10*
* `sd_j` = cell standard deviation vector, *default = 1.75*

> all vectors are lengeht 6, or recycled if shorter

```{r}
gen2x3 <- function(mu_j = 2,  
                   n_j  = 10, 
                   sd_j = 1.75){
  expand.grid(drug    = c("Antidepressant", 
                          "Placebo"),
              therapy = c("Psychoanalysis",
                          "Group Therapy",
                          "Behavior Modification")) %>% 
    dplyr::mutate(mean = mu_j) %>% 
    dplyr::mutate(panic = purrr::map(mean, 
                                     rnorm, 
                                     n = n_j, 
                                     sd = sd_j)) %>% 
    tidyr::unnest(panic) %>% 
    dplyr::mutate(id = row_number()) %>% 
    dplyr::group_by(drug, therapy) %>% 
    dplyr::mutate(rep = row_number()) %>% 
    dplyr::ungroup() %>% 
    dplyr::mutate(id = factor(id)) %>% 
    dplyr::mutate(reg = factor(rep)) %>% 
    dplyr::select(id, rep, drug, therapy, panic)
}
```

\clearpage

### Two-way Means Table

Cells contain: *M* (*S*D) [*n*]

```{r}
tab2x3 <- function(df){
  df %>%
    dplyr::group_by(therapy, drug) %>% 
    dplyr::summarise(m  = panic %>% mean() %>% round(2),
                     sd = panic %>% sd() %>% round(2),
                     n  = n()) %>% 
    dplyr::mutate(cell = glue::glue("{m} ({sd}) [{n}]")) %>% 
    dplyr::select(drug, therapy, cell) %>% 
    tidyr::pivot_wider(names_from = therapy,
                       values_from = cell) %>% 
    pander::pander()
}
```

### Two-way Means Plot

```{r}
plot2x3 <- function(df){
  df %>% 
    ggplot(aes(x = therapy,
               y = panic,
               group = drug,
               shape = drug,
               color = drug)) +
    stat_summary(position = position_dodge(width = .2)) +
    stat_summary(aes(linetype = drug),
                 position = position_dodge(width = .2),
                 fun.y = mean,
                 geom = "line") +
    theme_bw()
}
```


\clearpage

## Generate Datasets

### Set random seed

```{r}
set.seed(1)
```


### Datasets

```{r}
df_0 <- gen2x3() # all defaults
```


# No Effects

## Exploratory Data Analysis

```{r}
tibble::glimpse(df_0)
```


```{r}
tab2x3(df_0)
```

```{r}
plot2x3(df_0)
```


