---
title: "Psy/Educ 6600: Unit 2 Homework"
subtitle: "Groundwork for Inference"
author: "Your Name"
date: "Spring 2018"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
    number_sections: false
---


\clearpage

# Chapter 1. DATA PREPARATION



```{r, include=FALSE}
# set global chunk options...  
#  this changes the defaults so you don't have to repeat yourself
knitr::opts_chunk$set(comment     = NA,
                      echo        = TRUE, 
                      warning     = FALSE, 
                      message     = FALSE,
                      fig.align   = "center", # center all figures
                      fig.width   = 4,        # set default figure width to 4 inches
                      fig.height  = 3)        # set default figure height to 3 inches

```

## Load Packages

* Make sure the packages are **installed** *(Package tab)*

```{r}
library(tidyverse)    # Loads several very helpful 'tidy' packages
library(readxl)       # Read in Excel datasets
library(furniture)    # Nice tables (by our own Tyson Barrett)
library(psych)        # Lots of nice tid-bits
library(car)          # Companion to "Applied Regression"
```

## Import Data, Define Factors, and Compute New Variables

* Make sure the **dataset** is saved in the same *folder* as this file
* Make sure the that *folder* is the **working directory**

> NOTE: I added the second line to convert all the variables names to lower case.  I still kept the `F` as a capital letter at the end of the five factor variables.

```{r}
data_clean <- read_excel("Ihno_dataset.xls") %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(genderF = factor(gender, 
                                 levels = c(1, 2),
                                 labels = c("Female", 
                                            "Male"))) %>% 
  dplyr::mutate(majorF = factor(major, 
                                levels = c(1, 2, 3, 4,5),
                                labels = c("Psychology",
                                           "Premed",
                                           "Biology",
                                           "Sociology",
                                           "Economics"))) %>% 
  dplyr::mutate(reasonF = factor(reason,
                                 levels = c(1, 2, 3),
                                 labels = c("Program requirement",
                                            "Personal interest",
                                            "Advisor recommendation"))) %>% 
  dplyr::mutate(exp_condF = factor(exp_cond,
                                   levels = c(1, 2, 3, 4),
                                   labels = c("Easy",
                                              "Moderate",
                                              "Difficult",
                                              "Impossible"))) %>% 
  dplyr::mutate(coffeeF = factor(coffee,
                                 levels = c(0, 1),
                                 labels = c("Not a regular coffee drinker",
                                            "Regularly drinks coffee")))  %>% 
  dplyr::mutate(hr_base_bps = hr_base / 60) %>% 
  dplyr::mutate(anx_plus = rowsums(anx_base, anx_pre, anx_post)) %>% 
  dplyr::mutate(hr_avg = rowmeans(hr_base + hr_pre + hr_post)) %>% 
  dplyr::mutate(statDiff = statquiz - exp_sqz)
```

\clearpage


# Chapter 5. Intro to Hypothesis Testing: 1 Sample `z`-Test

## 5C-3. 1 Sample `z`-Test compared to historic controls for `mathquiz` and `statquiz`

**TEXTBOOK QUESTION:** *(A) In the past 10 years, previous stats classes who took the same math quiz that Inho's students took **averaged 28** with a **standard deviation of 8.5**. What is the two-tailed p value for Inho's students with respect to that past population? (Don't forget that the N for mathquiz is not 100.)  Would you say that Inho's class performed significantly better than previous classes? Explain. (B) Redo part a assuming that the same previous classes had also taken the same statquiz and **averaged 6.1** with a **standard deviation of 2.5**.*

**DIRECTIONS:** Find the mean (`M`) and sample size (`n`) for `mathquiz` and `statquiz` and then work the rest of the statistical test by hand in the printed homework packet.  

> **NOTE:** You may use the `furniture::table1()` funciton gives the mean, but it only gives the total `n` for all variables.  Since some students were missing the math quiz, but not the stat quiz the sample sizes are different.  So use the `psych::describe()` function to get the means and the sample size for each variable.

```{r}
# Find the mean and n for: mathquiz, statquiz

```








\clearpage

## 5C-4. Test for Normaity for `mathquiz` and `statquiz`

**TEXTBOOK QUESTION:** *Test both the math quiz and stat quiz variables for their resemblance to normal distributions. Based on skewness, kurtosis, and the Shapiro-Wilk statistic, which variable has a sample distribution that is not very consistent with the assumption of normality in the population?*

### Skewness and Kurtosis

**DIRECTIONS:** Find the skewness and kurtosis for `mathquiz` and `statquiz` 

> **NOTE:** Yes, you just did this above using the `psych::describe()` function... so you may skip it here if you want.

```{r, eval=FALSE}
# Find the skewness and kurtosis for: mathquiz, statquiz

```


### Shapiro-Wilk's Test

**DIRECTIONS:**  Use the `shapiro.test()` function to test for normality in a small'ish sample.

> **NOTE:** You must use a `dplyr::pull()` step to pull out one variable from the dataset before you can use the `shapiro.test()` function.

```{r}
# Shapiro-Wilk's Normality Test for: mathquiz

```

```{r}
# Shapiro-Wilk's Normality Test for: statquiz

```

\clearpage

### Create Histograms

**DIRECTIONS:**  Use `geom_histogram()` after setting the `ggplot(aes())`.  Make sure to try different `bins = #` or `binwidth = #` to get a 'good looking' plot.

> **NOTE:** For histograms, you do need to specify the variable name as `x`in the `aes(x = variable)` option.  

```{r}
# Histogram for: mathquiz

```

```{r}
# Histogram for: statquiz

```

\clearpage

### Create QQ Plots

**DIRECTIONS:**  Use `geom_qq()` after setting the `ggplot(aes())`.  

> **NOTE:** For qq plots, you do need to specify the variable name as `sample`in the `aes(sample = variable)` option.  

```{r}
# Histogram for: mathquiz

```

```{r}
# Histogram for: statquiz

```





\clearpage

# Chapter 6. Confidence Interval Estimation: The `t` Distribution

## 6C-1. 1-sample `t`-tests for `anx_base`, `anx_pre`, and `anx_post`


**TEXTBOOK QUESTION:** *Perform one-sample t tests to determine whether the baseline, pre-, or postquiz anxiety scores of Inho's students differ significantly ( $\alpha = .05$, two-tailed) from the mean ($\mu = 18$) found by a very large study of college students across the country. Find the 95% Cconfidence interval for the population mean for each of the **three** anxiety measures.*


**DIRECTIONS:**  Use the `t.test(mu = #)` function to perform a 1 sample `t`-test.  Make sure to sepify the Null hypothesis value for $\mu$.

> **NOTE:** You must use a `dplyr::pull()` step to pull out one variable from the dataset before you can use the `t.test()` function.

```{r}
# 1-sample t-test for: anx_base

```

```{r}
# 1-sample t-test for: anx_base

```

```{r}
# 1-sample t-test for: anx_base

```






\clearpage

## 6C-2. 1-sample `t`-tests for `hr_base` among MEN

**TEXTBOOK QUESTION:** *Perform a one-sample t test to determine whether the average baseline heart rate of Inho's **male** students differs significantly from the **mean** heart rate ($\mu = 70$) for college-aged men at the **.01 level**, two-tailed. Find the **99%** confidence intervals for the population mean represented by Inho's **male** students.*

**DIRECTIONS:** Similar to the last problem, use the `t.test(mu = #)` function to perform a 1 sample `t`-test.  This time, make sure the subset out the males only with a `dplyr::filter()` step prior to the `dplyr::pull()` step.

> **note:** To change from the default 95% confidence intervals, make sure to specify `conf.level = 0.99` inside the `t.test()` function.

```{r}
# 1-sample t-test for MALES: hr_base

```





\clearpage 

## 6C-3. 1-sample `t`-tests for `hr_post` among FEMALE

**TEXTBOOK QUESTION:** *Perform a one-sample t test to determine whether the average postquiz heart rate of Inho's **female** students differs significantly ($\alpha = .05$, two-tailed) from the **mean** resting heart rate ($\mu = 72$) for college-aged women. Find the 95% confidence interval for the population mean represented by Inho's **female** students.*

**DIRECTIONS:** This time, subset out WOMEN and choose the post-quiz heart rate.  Also, use a different population null value.

```{r}
# 1-sample t-test for MALES: hr_base

```







\clearpage

# Chapter 7. Independent Samples `t`-Test for Means

## 7C-1. Independent Samples `t`-Test for Mean `hr_base` by `genderF`

**TEXTBOOK QUESTION:**  *Perform a two-sample t test to determine whether there is a statistically significant difference in **baseline heart rate** between the **men and the women** of Inho's class. Do you have **homogeneity of variance**? Report your results as they might appear in a journal article. Include the 95% CI for this gender difference.*

### Assumtion Check: Homogeneity of Variance

**DIRECTIONS:** Before performing the test, check to see if the assumption of homogeneity of variance is met using **Levene's Test**.  For a independent samples `t`-test for means, the men and women need to have the same amount of spread (SD) in their baseline hear rates.

 > **NOTE:** Use the `car:leveneTest()` function to do this.  Inside the funtion you need to specify at least three things (sepearated by commas):

> * the formula:  `continuous_var ~ grouping_var` (replace with your variable names)

> * the dataset:  `data = .` to pipe it from above

> * the center:  `center = "mean"` since we are comparing means

```{r}

```


\clearpage

### Perform the `t`-Test for Means in 2 Indep Groups 

**DIRECTIONS:** Test if men and women have different baseline heart rates using the `t.test()` function.

>Use the same `t.test()` funtion we have used in the prior chapters.  This time you need to speficy a few more options:

> * the formula:  `continuous_var ~ grouping_var` (replace with your variable names)

> * the dataset:  `data = .` to pipe it from above

> * independent vs. paired:  `paired = FALSE` (this is the default)

> * is homogeneity satified: `var.equal = TRUE` (NOT the default)

> * confidence level: `conf.level = #` (defults to .95)


```{r}
# indep groups t-test for means: hr_base by genderF

```






\clearpage

## 7C-5. Independent Samples `t`-Test for Mean `hr_post` by `coffeeF`

**TEXTBOOK QUESTIONS:** *Perform a two-sample t test to determine whether **coffee drinkers** exhibited significantly higher **postquiz heart rates** than nondrinkers at the .05 level. Is this t test significant at the .01 level? Find the **99%** confidence interval for the difference of the two population means and explain its connection to your decision regarding the null hypothesis at the **.01 level**.*


### Assumtion Check: Homogeneity of Variance

**DIRECTIONS:** Just like the last question, run **Levene's test** first.

```{r}

```

### Perform the `t`-Test for Means in 2 Indep Groups 

**DIRECTIONS:** Make sure to change the confidence level to **99%**.

```{r}
# indep groups t-test for means: hr_post by coffeeF

```

