---
title: "Psy/Educ 6600: Chapter 2"
subtitle: "DISTRIBUTION and UNIVARIATE PLOTS"
author: "Your Name"
date: "Spring 2018"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
    number_sections: false
---

\clearpage

# PREPARATION

```{r, include=FALSE}
# set global chunk options...  
#  this changes the defaults so you don't have to repeat yourself
knitr::opts_chunk$set(comment     = NA,
                      echo        = TRUE, 
                      warning     = FALSE, 
                      message     = FALSE,
                      fig.align   = "center", # center all figures
                      fig.width   = 4,        # set default figure width to 4 inches
                      fig.height  = 3)        # set default figure height to 3 inches
```

## Packages

* Make sure the packages are **installed** *(Package tab)*

```{r}
library(tidyverse)    # Loads several very helpful 'tidy' packages
library(readxl)       # Read in Excel datasets
library(furniture)    # Nice tables (by our own Tyson Barrett)
library(psych)        # Lots of nice tid-bits
```

\clearpage



# SECTION C

## Import Data, Define Factors, and Compute New Variables

* Make sure the **dataset** is saved in the same *folder* as this file
* Make sure the that *folder* is the **working directory**

> NOTE: I added the second line to convert all the variables names to lower case.  I still kept the `F` as a capital letter at the end of the five factor variables.

```{r}
data_clean <- read_excel("Ihno_dataset.xls") %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(genderF = factor(gender, 
                                 levels = c(1, 2),
                                 labels = c("Female", 
                                            "Male"))) %>% 
  dplyr::mutate(majorF = factor(major, 
                                levels = c(1, 2, 3, 4,5),
                                labels = c("Psychology",
                                           "Premed",
                                           "Biology",
                                           "Sociology",
                                           "Economics"))) %>% 
  dplyr::mutate(reasonF = factor(reason,
                                 levels = c(1, 2, 3),
                                 labels = c("Program requirement",
                                            "Personal interest",
                                            "Advisor recommendation"))) %>% 
  dplyr::mutate(exp_condF = factor(exp_cond,
                                   levels = c(1, 2, 3, 4),
                                   labels = c("Easy",
                                              "Moderate",
                                              "Difficult",
                                              "Impossible"))) %>% 
  dplyr::mutate(coffeeF = factor(coffee,
                                 levels = c(0, 1),
                                 labels = c("Not a regular coffee drinker",
                                            "Regularly drinks coffee")))  %>% 
  dplyr::mutate(hr_base_bps = hr_base / 60) %>% 
  dplyr::mutate(anx_plus = rowsums(anx_base, anx_pre, anx_post)) %>% 
  dplyr::mutate(hr_avg = rowmeans(hr_base + hr_pre + hr_post)) %>% 
  dplyr::mutate(statDiff = statquiz - exp_sqz)
```

\clearpage



## Question C-1. Frequency Distribution and Bar Chart

Request a frequency distribution using the `furniture::tableF(continuous_var)` function  

```{r}
# Frequency distrubution: majorF

```



Create a bar chart using `geom_bar()` for the Undergraduate Major (`majorF`) variable for Ihno's students.

> Make sure to add the variable of interest into the asthetics: `ggplot(aes(continuous_var))` before adding the `geom_bar()` layer.

```{r}
# Bar Plot: majorF

```

\clearpage



## Question C-2. Bar Charts

Repeat Exercise 1 for the variables `prevmath` and `phobia`. 

> IN THE WRITEUP: Would it make sense to request a histogram instead of a bar chart for phobia ? Discuss.

```{r}
# Frequency distrubution: prevmath

```



```{r}
# Bar Plot: prevmath

```



```{r}
# Frequency distrubution: phobia


```



```{r}
# Bar Plot: phobia


```

\clearpage



## Question C-3. Frequency Distribution and Histogram

Request a frequency distribution and a histogram for the variable `statquiz`.  Use the option in the function `geom_histogram(bins = #)` to change the number of bins or `geom_histogram(binwidth = #)` to change the bin width to give a better figure.

> IN THE WRITEUP: Describe the shape of this distribution.

```{r}
# Frequency distrubution: statquiz

```



```{r}
# Histogram: statquiz, with a different number/width of bins

```

\clearpage



## Question C-4. Frequency Distribution and Histogram

Request a frequency distribution and a histogram for the variables baseline anxiety (`anx_base`) and baseline heart rate (`hr_base`). 

> IN THE WRITEUP: Comment on R's choice of class intervals for each histogram.

```{r}
# Frequency distrubution: anx_base

```



```{r, fig.height=2.5}
# Histogram: anx_base

```

\clearpage



```{r}
# Frequency distrubution: hr_base

```



```{r}
# Histogram: hr_base

```

\clearpage



## Question C-6.  Histograms -by- a Factor

Request Histograms for the variables `anx_base` and `hr_base` divided by `genderF` using an additional `facet_grid(group_var ~ .)` layer to create two plots.

```{r}
# Histogram: anx_base, by genderF

```



```{r}
# Histogram: hr_base, by genderF

```

\clearpage



## Question C-9. Deciles and Quartiles

Using the `quantile(probs = c(#, #, ..., #))` function, request the deciles and quartiles for the `phobia` variable.  

> Make sure to add a `dplyr::pull(varname)` step to pull out only the one variable you are interested in.

```{r}
# Deciles: phobia

```



```{r}
# Quartiles: phobia

```



## Question C-10. Various Percentiles

Request the following percentiles for the variables `hr_base` and `hr_pre`: 15, 30, 42.5, 81, and 96.

```{r}
# Percentiles: hr_base

```



```{r}
# Percentiles: hr_pre

```


