---
title: "Psy/Educ 6600: Chapter 4"
subtitle: "STANDARDIZED SCORES"
author: "Your Name"
date: "Spring 2018"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
    number_sections: false
---

\clearpage

# PREPARATION

```{r, include=FALSE}
# set global chunk options...  
#  this changes the defaults so you don't have to repeat yourself
knitr::opts_chunk$set(comment     = NA,
                      echo        = TRUE, 
                      warning     = FALSE, 
                      message     = FALSE,
                      fig.align   = "center", # center all figures
                      fig.width   = 4,        # set default figure width to 4 inches
                      fig.height  = 3)        # set default figure height to 3 inches
```

## Packages

* Make sure the packages are **installed** *(Package tab)*

```{r}
library(tidyverse)    # Loads several very helpful 'tidy' packages
library(readxl)       # Read in Excel datasets
library(furniture)    # Nice tables (by our own Tyson Barrett)
library(psych)        # Lots of nice tid-bits
```

\clearpage



# SECTION C

## Import Data, Define Factors, and Compute New Variables

* Make sure the **dataset** is saved in the same *folder* as this file
* Make sure the that *folder* is the **working directory**

> NOTE: I added the second line to convert all the variables names to lower case.  I still kept the `F` as a capital letter at the end of the five factor variables.

```{r}
data_clean <- read_excel("Ihno_dataset.xls") %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(genderF = factor(gender, 
                                 levels = c(1, 2),
                                 labels = c("Female", 
                                            "Male"))) %>% 
  dplyr::mutate(majorF = factor(major, 
                                levels = c(1, 2, 3, 4,5),
                                labels = c("Psychology",
                                           "Premed",
                                           "Biology",
                                           "Sociology",
                                           "Economics"))) %>% 
  dplyr::mutate(reasonF = factor(reason,
                                 levels = c(1, 2, 3),
                                 labels = c("Program requirement",
                                            "Personal interest",
                                            "Advisor recommendation"))) %>% 
  dplyr::mutate(exp_condF = factor(exp_cond,
                                   levels = c(1, 2, 3, 4),
                                   labels = c("Easy",
                                              "Moderate",
                                              "Difficult",
                                              "Impossible"))) %>% 
  dplyr::mutate(coffeeF = factor(coffee,
                                 levels = c(0, 1),
                                 labels = c("Not a regular coffee drinker",
                                            "Regularly drinks coffee")))  %>% 
  dplyr::mutate(hr_base_bps = hr_base / 60) %>% 
  dplyr::mutate(anx_plus = rowsums(anx_base, anx_pre, anx_post)) %>% 
  dplyr::mutate(hr_avg = rowmeans(hr_base + hr_pre + hr_post)) %>% 
  dplyr::mutate(statDiff = statquiz - exp_sqz)
```

\clearpage



## Question C-1. Calculate z-Scores

Use the `dplyr::mutate(new_zscore_var = scale(old_orig_var))` function to create two new variables consisting of the *z scores* for the **anxiety** and **heart rate** measures at **baseline** in Ihno's data set. 

Then request *means* and *SD's* of the *z-score* variables with the `furniture::table1()` function to demonstrate that the means and SD s are `0` and `1`, respectively, in each case.

```{r}
# Descriptive Stats: baseline anx & hr, original and z-scores

```


