---
title: "Psy/Educ 6600: Chapter 6"
subtitle: "Confidence Interval Estimation: The t Distribution"
author: "Dr. Sarah Schwartz - Answer Key"
date: "Spring 2019"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
    number_sections: false
---

\clearpage

# PREPARATION

```{r, include=FALSE}
# set global chunk options...  
#  this changes the defaults so you don't have to repeat yourself
knitr::opts_chunk$set(comment     = NA,
                      echo        = TRUE, 
                      warning     = FALSE, 
                      message     = FALSE,
                      fig.align   = "center", # center all figures
                      fig.width   = 4,        # set default figure width to 4 inches
                      fig.height  = 3)        # set default figure height to 3 inches

```

## Packages

* Make sure the packages are **installed** *(Package tab)*

```{r}
library(tidyverse)    # Loads several very helpful 'tidy' packages
library(readxl)       # Read in Excel datasets
library(furniture)    # Nice tables (by our own Tyson Barrett)
library(psych)        # Lots of nice tid-bits
library(car)          # Companion to "Applied Regression"
```

\clearpage



# SECTION C

## Import Data, Define Factors, and Compute New Variables

* Make sure the **dataset** is saved in the same *folder* as this file
* Make sure the that *folder* is the **working directory**

> NOTE: I added the second line to convert all the variables names to lower case.  I still kept the `F` as a capital letter at the end of the five factor variables.

```{r}
data_clean <- read_excel("Ihno_dataset.xls") %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(genderF = factor(gender, 
                                 levels = c(1, 2),
                                 labels = c("Female", 
                                            "Male"))) %>% 
  dplyr::mutate(majorF = factor(major, 
                                levels = c(1, 2, 3, 4,5),
                                labels = c("Psychology",
                                           "Premed",
                                           "Biology",
                                           "Sociology",
                                           "Economics"))) %>% 
  dplyr::mutate(reasonF = factor(reason,
                                 levels = c(1, 2, 3),
                                 labels = c("Program requirement",
                                            "Personal interest",
                                            "Advisor recommendation"))) %>% 
  dplyr::mutate(exp_condF = factor(exp_cond,
                                   levels = c(1, 2, 3, 4),
                                   labels = c("Easy",
                                              "Moderate",
                                              "Difficult",
                                              "Impossible"))) %>% 
  dplyr::mutate(coffeeF = factor(coffee,
                                 levels = c(0, 1),
                                 labels = c("Not a regular coffee drinker",
                                            "Regularly drinks coffee")))  %>% 
  dplyr::mutate(hr_base_bps = hr_base / 60) %>% 
  dplyr::mutate(anx_plus = rowsums(anx_base, anx_pre, anx_post)) %>% 
  dplyr::mutate(hr_avg = rowmeans(hr_base + hr_pre + hr_post)) %>% 
  dplyr::mutate(statDiff = statquiz - exp_sqz)
```

\clearpage



## Question C-1. 1-sample `t`-tests for `anx_base`, `anx_pre`, and `anx_post`

**TEXTBOOK QUESTION:** *Perform one-sample t tests to determine whether the baseline, pre-, or postquiz anxiety scores of Ihno’s students differ significantly ( $\alpha = .05$, two-tailed) from the mean ($\mu = 18$) found by a very large study of college students across the country. Find the 95% Cconfidence interval for the population mean for each of the **three** anxiety measures.*

----------------------------

**DIRECTIONS:**  Use the `t.test(mu = #)` function to perform a 1 sample `t`-test.  Make sure to specify the Null hypothesis value for $\mu$.

> **NOTE:** You must use a `dplyr::pull()` step to pull out one variable from the dataset before you can use the `t.test()` function.

```{r}
# 1-sample t-test for: anx_base

```

---------------------------

```{r}
# 1-sample t-test for: anx_pre

```

\clearpage



```{r}
# 1-sample t-test for: anx_post

```

\clearpage



## Question C-2. 1-sample `t`-tests for `hr_base` among MEN

**TEXTBOOK QUESTION:** *Perform a one-sample t test to determine whether the average baseline heart rate of Ihno’s **male** students differs significantly from the **mean** heart rate ($\mu = 70$) for college-aged men at the **.01 level**, two-tailed. Find the **99%** confidence intervals for the population mean represented by Ihno’s **male** students.*

------------------------

**DIRECTIONS:** Similar to the last problem, use the `t.test(mu = #)` function to perform a 1 sample `t`-test.  This time, make sure the subset out the MEN only (`genderF == "Male"`) with a `dplyr::filter()` step prior to the `dplyr::pull()` step.

> **Note:** To change from the default 95% confidence intervals, make sure to specify `conf.level = 0.99` inside the `t.test()` function.

```{r}
# 1-sample t-test for MALES: hr_base

```


\clearpage 



## Question C-3. 1-sample `t`-tests for `hr_post` among FEMALE

**TEXTBOOK QUESTION:** *Perform a one-sample t test to determine whether the average postquiz heart rate of Ihno’s **female** students differs significantly ($\alpha = .05$, two-tailed) from the **mean** resting heart rate ($\mu = 72$) for college-aged women. Find the 95% confidence interval for the population mean represented by Ihno’s **female** students.*

----------------------------------

**DIRECTIONS:** This time, subset out WOMEN (`genderF == "Female"`) and choose the post-quiz heart rate.  Also, use a different population null value ($\mu$).

```{r}
# 1-sample t-test for FEMALES: hr_post

```


